---
title: "Exploring the Relationship between the Mode of Transmission and the Mileage"
author: "Jeevithiesh Duggani"
date: "08/05/2020"
output: pdf_document
---
```{r setup,echo=FALSE,eval=TRUE}
    options(scipen = 10)
```

## Executive Summary
From the dataset `mtcars`,  this project explores the relationship between the `mode of transmission` and `mileage in miles per gallon (MPG)`. The project aims to answer the following two questions.

1. Is an automatic or manual transmission better for MPG
2. Quantify the MPG difference between automatic and manual transmissions

## Import libraries
Import the necessary libraries
```{r Import libraries,results='hide'}
    library(ggplot2)
```

## Initialize the data
Import the dataset `mtcars`. 
```{r Import mtcars}
    data("mtcars")
```

## Exploratory Data Analysis
Consider the average `mpg` for each level of `am`.
```{r avg mpg per level of am}
    aggregate(mpg ~ am, mtcars, mean)
```

It can be seen from the above data frame that the `average mileage` is **better** for `manual transmission` than for `automatic transmission`.

## Regression Analysis

> Hypothesis : The value of mpg for manual transmission is greater than that for automatic transmission

Find the `coefficients` for the `linear regression` with only `mpg` as outcome and `am` as predictor.
```{r coefficients of linear regression}
    fit <- lm(mpg ~ am, data = mtcars)
    coefficients(summary(fit))
```

Find the confidence intervals for `am`.
```{r confidence intervals}
    confint(fit)["am",]
```

The entire `confidence interval` lies above `0` and the `p-value` (`r summary(fit)$coefficients[2,4]`) is less than `0.05`. 

> Therfore the hypothesis is not to be rejected.

The value of R squared is `r round(summary(fit)$r.square, 4)`, i.e. only `r paste0(round(summary(fit)$r.square*100,4),"%")` of the variance of `mpg` is caused by `am`.

Perform Analysis of Variance over the data.
```{r perform analysis of variance}
    analyseVariance <- aov(mpg ~ ., mtcars)
    summary(analyseVariance)
```

The terms with `p-value` less than `0.05` are `cyl`, `disp`, `wt`.

Find the `linear regression` with `mpg` as outcome and `cyl`, `wt`, `disp` and `am` as predictors.

```{r coefficients of multiple linear regression}
    fitMult <- lm(mpg ~ cyl + wt + disp + am, data = mtcars)
    coefficients(summary(fitMult))
```

The R squared is `r round(summary(fitMult)$r.square, 4)` which is sufficiently large.

> The p values of cyl and wt are below 0.05, suggesting that these are the confunding values between mpg and am

Identify the coefficients after removing `disp` as a predictor
```{r coefficients of final linear regression}
    fitFin <- lm(mpg ~ cyl + wt + am, data = mtcars)
    coefficients(summary(fitFin))
```

The `coefficient` for `am` is `0.1765`.

> The value of mpg increases by 0.1765 when the mode of transmission changes from automatic to manual.

## Appendix
Plot the graph between `mpg` and `am`
```{r plot graph mpg vs am}
    mtcars$am <- as.factor(mtcars$am)
    levels(mtcars$am) <- c("Automatic", "Manual")
    graph <- ggplot(mtcars, aes(x = factor(am), y = mpg))
    graph <- graph + geom_boxplot(aes(fill = factor(am)))
    graph <- graph + labs(title = "MPG for each Transmission", 
                          x = "Transmission", y = "Mileage")
    graph
```

From the above analysis, it can be initially observed that `manual transmission` is indeed **better** for `mileage`.

Plot for fit.
```{r plot graph for fit}
    par(mfrow = c(2,2))
    plot(fit)
```

plot for fitMult.
```{r plot graph for fitMult}
    par(mfrow = c(2,2))
    plot(fitMult)
```

plot for fitFin.
```{r plot graph for fitFin}
    par(mfrow = c(2,2))
    plot(fitFin)
```

The `residuals vs fitted` graphs of `fitMult` and `fitFin` are very similar, indicating that the chosen confunding variables are the only confounding variables.

                                    **End**